# The Spaghetti Code Challenge

## Background

The goal of this challenge is to replicate Figure 3 of Schaefer et al. 2000, as shown in README.md.

## Data

In the aforementioned figure, stream contents from four different sites were analyzed: three sites around Quebrada Cuenca and one in Rio Meyes, Puente Roto. Each site corresponds to the following assigned variable.

cuenca1 = Quebrada Cuenca Site 1 (Q1) cuenca2 = Quebrada Cuenca Site 2 (Q2) cuenca3 = Quebrada Cuenca Site 3 (Q3) puente_roto = Rio Mameyes, Puente Roto (PRM)

```{r}
library(tidyverse)

cuenca1 <- read_csv("../data/QuebradaCuenca1-Bisley.csv")
cuenca2 <- read_csv("../data/QuebradaCuenca2-Bisley.csv")
cuenca3 <- read_csv("../data/QuebradaCuenca3-Bisley.csv")
puente_roto <- read_csv("../data/RioMameyesPuenteRoto.csv")
```

## Methods

To process each dataset, I created a function called "moving-average". This function takes the mean concentration of each ion of interest (Nitrate(NO3-N), Ammonium (NH4-N), Potassium (K), Magnesium (Mg), and Calcium (Ca)) every nine weeks thoughout the sampling period. The function subsequently generates new datasets that start with "mov-avg" containing the nine-week average.

```{r}
# Source moving-average function
source("../R/moving-average.R")

# Apply function to each datset

mov_avg_cuenca1 <- moving_average(cuenca1)
mov_avg_cuenca2 <- moving_average(cuenca2)
mov_avg_cuenca3 <- moving_average(cuenca3)
mov_avg_roto <- moving_average(puente_roto)
```

To visuallize data from all four streams at the same time, I bind the datasets together and pivot the columns. This allows me to plot concentration over time for each ion of interest.

```{r}
# Now bind data
bind_streams <- bind_rows(
  mov_avg_cuenca1,
  mov_avg_cuenca2,
  mov_avg_cuenca3,
  mov_avg_roto
)

# Pivot data in cuenca1 for visuallization
pivot_streams <- bind_streams |>
  pivot_longer(
    cols = c(`NO3-N`, `NH4-N`, K, Mg, Ca),
    names_to = "Ions",
    values_to = "Concentration"
  )
```

## Results

\[Insert results\]

```{r}
ggplot(
  data = pivot_streams,
  mapping = aes(
    x = window_start,
    y = Concentration,
    color = Sample_ID
  )
) +
  geom_line() +
  theme_classic() +
  labs(
    title = "Ion Concentration over Time",
    x = "Sampling Dates",
    y = "Concentration of Ion (mg/l)"
  ) +
  facet_wrap(~Ions, nrow = 5, ncol = 1, scales = "free")
```